---
title: "Project Proposal"
format: html
author:
  - "Aleksandar Neshov"
  - "Dinh Marcus Nguyen"
  - "Alexander Toropov"
date: today
---

## Dataset

### Source

<https://opendata.muenchen.de/dataset/monatszahlen-tourismus>

### License

Datenlizenz Deutschland Namensnennung 2.0 (dl-by-de)

### Description

The dataset contains monthly counts of guests and overnight stays in Munich, split by domestic and international visitors.

### The key variables are:

-   MONATSZAHL: metric type (guests or overnight stays) chr

-   AUSPRAEGUNG: visitor origin (domestic / international), chr

-   JAHR: year of observation, dbl

-   MONAT: month of observation, chr

-   WERT: the monthly count, dbl

### Dimensions:

-   1632 rows

-   10 columns

## Research Questions

1.  How do monthly tourism patterns in Munich vary across the year, and which months show the highest and lowest tourism activity?

    Claim type: Descriptive / exploratory. We describe seasonal patterns and identify months with high or low tourism activity.

2.  How did Munich tourism change during the COVID-19 period, and how did domestic and international tourism recover afterwards?

    Claim type: Descriptive / exploratory. We describe visible changes during and after COVID-19, without making strong causal claims.

3.  How do domestic and international tourism differ in their long-term trends and seasonal patterns?

    Claim type: Exploratory. We compare Inland and Ausland to see whether their trends and seasonal patterns differ.

## Context and target audience

Our project is framed in a journalistic / public-interest context. We want to explore how tourism in Munich has changed over time, especially seasonal patterns, the COVID-19 period, and differences between domestic and international tourism.

The target audience is a general audience interested in Munich, such as local residents, students, journalists, city administration, and tourism-related businesses. The analysis should therefore be understandable without advanced statistical knowledge.

Suitable presentation formats are: - a short data report - a Quarto document with explanatory text and visualisations - line charts showing trends over time - seasonal plots comparing months - summary tables comparing domestic and international tourism

## Initial EDA

```{r}
library(tidyverse)
library(janitor)
library(skimr)

tourism <- read_csv("data/raw/tourism.csv")

glimpse(tourism)
```

### Summary statistics for key variables

```{r}
tourism |>
  summarise(
    rows = n(),
    columns = ncol(tourism),
    min_year = min(JAHR),
    max_year = max(JAHR),
    missing_wert = sum(is.na(WERT)),
    min_wert = min(WERT, na.rm = TRUE),
    median_wert = median(WERT, na.rm = TRUE),
    mean_wert = mean(WERT, na.rm = TRUE),
    max_wert = max(WERT, na.rm = TRUE)
  )
```

The dataset contains 1,632 rows and 10 columns. The years range from 2006 to 2026. The main outcome variable WERT has 72 missing values. For non-missing values, WERT ranges from 4,785 to 19,689,217, with a median of 498,710 and a mean of about 1,014,664.

```{r}
tourism |>
  count(MONATSZAHL)

tourism |>
  count(AUSPRAEGUNG)
```

The variable MONATSZAHL contains two tourism measures: Gäste and Übernachtungen, each with 816 observations. The variable AUSPRAEGUNG contains three categories: Ausland, Inland, and insgesamt, each with 544 observations.

```{r}
tourism |>
  summarise(across(everything(), ~ sum(is.na(.))))
```

WERT: 72 missing values

VERAEND_VORMONAT_PROZENT: 192 missing values

VERAEND_VORJAHRESMONAT_PROZENT: 72 missing values

ZWOELF_MONATE_MITTELWERT: 192 missing values

### Plot 1: Distribution of tourism values

```{r}
tourism %>% 
  ggplot(aes(x = WERT)) +
  geom_histogram(bins = 30) +
  scale_x_log10() +
  labs(
    title = "Distribution of tourism values",
    x = "Tourism value WERT, log scale",
    y = "Number of observations"
  )
```

The distribution of WERT is strongly right-skewed. Most observations are much smaller than the largest values. The highest values likely come from yearly totals where MONAT is "Summe", while regular monthly observations are lower. A log scale makes the distribution easier to inspect.

### Plot 2: Tourism trend over time

```{r}
tourism %>% 
  filter(AUSPRAEGUNG %in% c("Inland", "Ausland")) %>% 
  group_by(JAHR, MONATSZAHL, AUSPRAEGUNG) %>% 
  summarise(total_wert = sum(WERT, na.rm = TRUE), .groups = "drop") %>% 
  ggplot(aes(x = JAHR, y = total_wert, color = AUSPRAEGUNG)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ MONATSZAHL, scales = "free_y") +
  labs(
    title = "Tourism in Munich over time",
    x = "Year",
    y = "Total tourism value",
    color = "Origin category"
  )
```

This plot gives a first overview of tourism trends over time. It compares domestic and international tourism and separates guests from overnight stays. The plot is relevant for our research questions about long-term trends and changes around the COVID-19 period.

### Plot 3: Monthly pattern

```{r}
tourism %>% 
  filter(MONAT != "Summe") %>% 
  filter(AUSPRAEGUNG %in% c("Inland", "Ausland")) %>% 
  mutate(
    month = as.integer(str_sub(MONAT, 5, 6))
  ) %>% 
  group_by(month, MONATSZAHL, AUSPRAEGUNG) %>% 
  summarise(mean_wert = mean(WERT, na.rm = TRUE), .groups = "drop") %>% 
  ggplot(aes(x = month, y = mean_wert, color = AUSPRAEGUNG)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ MONATSZAHL, scales = "free_y") +
  scale_x_continuous(breaks = 1:12) +
  labs(
    title = "Average monthly tourism patterns in Munich",
    x = "Month",
    y = "Average tourism value",
    color = "Origin category"
  )
```

This plot helps us examine seasonal patterns. It shows average tourism values by month and compares domestic and international tourism. This is relevant for identifying months with particularly high or low tourism activity.

## Notes on data quality issues

Several data quality issues are already visible in the initial exploration. First, the variable MONAT contains both monthly values such as "202401" and yearly totals marked as "Summe". These should not be mixed without care, because yearly totals are much larger than monthly observations.

Second, WERT has 72 missing values. These missing values appear especially relevant for 2026, which is likely incomplete at the time of data publication.

Third, the variables VERAEND_VORMONAT_PROZENT and ZWOELF_MONATE_MITTELWERT have more missing values than WERT. This is expected because percentage changes and moving averages cannot always be calculated for early or incomplete observations.

Fourth, AUSPRAEGUNG includes "insgesamt", which is the total of "Inland" and "Ausland". Therefore, "insgesamt" should not be treated as an independent third group when comparing domestic and international tourism.

Finally, MONAT is stored as a character variable. For seasonal analysis, we will likely need to extract the month number from values such as "202401".

## Analysis Plan

### Research Question 1:

How do monthly tourism patterns in Munich vary across the year, and which months show the highest and lowest tourism activity?

### Variables used:

WERT = tourism value

MONAT = month / time variable

MONATSZAHL = tourism measure: Gäste or Übernachtungen

AUSPRAEGUNG = origin category: Inland, Ausland, insgesamt

### Analytical approach:

We will use descriptive statistics and visualisations. We plan to calculate average tourism values for each month and compare seasonal patterns between Gäste and Übernachtungen. Line plots or bar charts by month are appropriate because the question focuses on seasonal patterns across the year.

### Cleaning / transformation steps:

We will need to extract the month number from MONAT, because values such as 202401 include both year and month. Rows where MONAT = "Summe" will need to be handled separately because they are yearly totals, not monthly observations. For this research question, we will likely exclude Summe rows.

### Uncertainties / risks:

The main risk is mixing monthly values with yearly totals. Another issue is that missing values, especially in incomplete years such as 2026, could distort average monthly patterns. We may need to exclude incomplete years or missing WERT values.

## Research Question 2:

How did Munich tourism change during the COVID-19 period, and how did domestic and international tourism recover afterwards?

### Variables used:

WERT = tourism value

JAHR = year

MONATSZAHL = Gäste or Übernachtungen

AUSPRAEGUNG = Inland, Ausland, insgesamt

### Analytical approach:

We will use time trend analysis with line plots and summary tables. We plan to compare tourism values before, during, and after the COVID-19 period. This can be done by comparing years such as pre-2020, 2020–2021, and post-2021. The approach is descriptive and exploratory, not fully causal.

### Cleaning / transformation steps:

We will need to decide whether to use monthly observations or yearly totals. For an annual trend analysis, we can either use MONAT = "Summe" or calculate yearly totals from monthly rows. We should not mix both. We will also need to handle missing values and possibly exclude 2026 because it is incomplete.

### Uncertainties / risks:

The main limitation is that the dataset alone cannot prove that COVID-19 caused the observed changes. It can only show changes during the COVID-19 period. Other factors, such as travel restrictions, economic conditions, or event cancellations, are not included as variables. International and domestic tourism may also be affected by different external factors.

## Research Question 3:

How do domestic and international tourism differ in their long-term trends and seasonal patterns?

### Variables used:

WERT = tourism value

JAHR = year

MONAT = month

AUSPRAEGUNG = Inland and Ausland

MONATSZAHL = Gäste and Übernachtungen

### Analytical approach:

We will compare Inland and Ausland using grouped line plots, seasonal plots, and summary statistics. This allows us to see whether domestic and international tourism follow similar or different patterns over time and across months.

### Cleaning / transformation steps:

For this question, we will focus on AUSPRAEGUNG = Inland and AUSPRAEGUNG = Ausland. The category insgesamt should not be treated as an independent third group because it is the total of domestic and international tourism. We will also extract the month number from MONAT for seasonal analysis and handle missing WERT values.

### Uncertainties / risks:

The main risk is interpreting differences too strongly. The dataset shows patterns, but it does not include reasons for those patterns. For example, changes in international tourism could be related to travel restrictions, global economic conditions, flight availability, or major events in Munich, but these variables are not included in the dataset.

## Workflow organisation

### Code style

We will follow the tidyverse style guide to keep the code readable and consistent.

Main conventions:

-   use snake_case for object and variable names
-   use clear and meaningful names
-   use the tidyverse pipe %\>%
-   keep code chunks short and focused
-   add comments only where they explain non-obvious steps

To enforce this, we will use a documented convention in the project README. We may also use styler to format code automatically:

```{r}
install.packages("styler")
styler::style_file("proposal.qmd")
```

### Packages

We plan to use the following R packages:

tidyverse = data cleaning, summaries, and visualisation

readr = reading CSV files dplyr = data manipulation

ggplot2 = plots

lubridate = date/month handling

janitor = cleaning column names

broom = tidying model outputs, if we use a model

We may use renv to make the project reproducible across different computers:

```{r}
install.packages("renv")
renv::init()
renv::snapshot()
```

This would lock the package versions used in the project.

### Git workflow

We will use a simple task-based Git workflow.

The main branch will contain the stable version of the project. Each group member will work on a separate branch or on a task-specific branch.

Possible branches:

main proposal data-cleaning visualisations analysis-plan final-report

Possible task division:

Person 1: dataset description and research questions Person 2: initial data exploration and summary statistics Person 3: plots and visualisations Person 4: analysis plan, workflow section, and final Quarto formatting

Before merging into main, we will check that the Quarto document still renders correctly and that the code runs without errors.

### .gitignore

Our `.gitignore` file was generated from the standard templates for R, Python, Windows, macOS, and Linux, and extended with Quarto-specific entries.

We exclude operating-system files such as `.DS_Store`, `Thumbs.db`, and Linux temporary files. We also exclude R session files such as `.Rhistory`, `.RData`, and `.Rproj.user/`, because they are user-specific and should not be shared through Git.

For reproducibility and security, we exclude local environment files such as `.env`, `.Renviron`, virtual environments, cache folders, and temporary build files. Since the project uses Quarto, we also exclude generated Quarto folders and cache files such as `/.quarto/`, `/_site/`, `/*_files/`, and `/*_cache/`.

Rendered output files and cached intermediate files should be regenerated from the Quarto source document instead of being manually tracked in Git.
