---
title: "Analysis"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
author:
  - "Aleksandar Neshov"
  - "Dinh Marcus Nguyen"
  - "Alexander Toropov"
date: today
---

## Research Question 1

**How do monthly tourism patterns in Munich vary across the year, and which months show the highest and lowest tourism activity?**

This is a descriptive analysis of Munich's seasonal tourism cycle. Guest arrivals and overnight stays are examined separately because they measure different aspects of tourism: arrivals count the number of visitors checking into accommodation, while overnight stays measure the total number of nights spent there. Domestic and international tourism are also separated to determine whether both visitor groups follow the same seasonal calendar.

The main description of the usual seasonal pattern excludes 2020 and 2021 because these exceptional years could shift the monthly averages away from normal conditions. The interactive plot still includes the COVID-only and all-years averages so that this decision can be assessed visually rather than simply hiding those observations.

```{r setup-rq1}
#| echo: false
#| warning: false
#| message: false

library(dplyr)
library(ggplot2)
library(plotly)
library(tsibble)
library(gt)

toursim_data <- readRDS("data/processed/tourism_clean.rds")
```

## Data preparation

Three versions are calculated:

- `seasonal_data` uses all available years
- `seasonal_data_no_covid` excludes 2020 and 2021 and represents the usual seasonal baseline
- `seasonal_data_only_covid` uses only 2020 and 2021

```{r prepare-seasonal-data-rq1}
summarise_seasonal_data <- function(data) {
  data %>%
    filter(AUSPRAEGUNG %in% c("Inland", "Ausland")) %>%
    mutate(month = as.integer(format(MONAT, format = "%m"))) %>%
    group_by(month, MONATSZAHL, AUSPRAEGUNG) %>%
    summarise(
      mean_wert = mean(WERT, na.rm = TRUE),
      .groups = "drop"
    )
}

seasonal_data <- toursim_data %>%
  summarise_seasonal_data()

seasonal_data_no_covid <- toursim_data %>%
  filter(!JAHR %in% c(2020, 2021)) %>%
  summarise_seasonal_data()

seasonal_data_only_covid <- toursim_data %>%
  filter(JAHR %in% c(2020, 2021)) %>%
  summarise_seasonal_data()
```

## Interactive seasonal comparison

The plot contains separate panels for guest arrivals and overnight stays because their values have different magnitudes. Each panel has six lines: domestic and international tourism for the three periods. The all-years lines show the overall averages, the non-COVID lines show the usual pattern, and the COVID-only lines show how strongly 2020 and 2021 differed from it.

Hovering over a line reports the month, average value, and percentage difference from the corresponding non-COVID baseline. Legend entries can be clicked to hide lines, which makes it easier to compare only the periods or origins of interest.

```{r interactive-covid-comparison-rq1}
baseline_data <- seasonal_data_no_covid %>%
  select(
    month,
    MONATSZAHL,
    AUSPRAEGUNG,
    baseline_wert = mean_wert
  )

plot_data <- bind_rows(
  "All years" = seasonal_data,
  "Without COVID-19" = seasonal_data_no_covid,
  "COVID-19 only" = seasonal_data_only_covid,
  .id = "period"
) %>%
  left_join(
    baseline_data,
    by = c("month", "MONATSZAHL", "AUSPRAEGUNG")
  ) %>%
  mutate(
    period = factor(
      period,
      levels = c(
        "All years",
        "Without COVID-19",
        "COVID-19 only"
      )
    ),
    measure = factor(
      recode(
        MONATSZAHL,
        "Gäste" = "Guests",
        "Übernachtungen" = "Overnight stays"
      ),
      levels = c("Guests", "Overnight stays")
    ),
    origin = recode(
      AUSPRAEGUNG,
      "Inland" = "Domestic",
      "Ausland" = "International"
    ),
    series = paste(origin, period, sep = " — "),
    change_from_baseline = 100 * (mean_wert / baseline_wert - 1)
  )

covid_comparison_plot <- ggplot(
  plot_data,
  aes(
    x = month,
    y = mean_wert,
    color = series,
    group = series,
    text = paste0(
      "Measure: ", measure,
      "<br>Origin: ", origin,
      "<br>Period: ", period,
      "<br>Month: ", month.name[month],
      "<br>Average: ", scales::comma(round(mean_wert)),
      "<br>Difference from non-COVID baseline: ",
      scales::percent(change_from_baseline / 100, accuracy = 0.1)
    )
  )
) +
  geom_line(linewidth = 0.9) +
  facet_wrap(
    ~measure,
    scales = "free_y",
    ncol = 2
  ) +
  scale_x_continuous(
    breaks = 1:12,
    labels = month.abb,
    expand = expansion(mult = c(0.01, 0.01))
  ) +
  scale_y_continuous(
    labels = scales::label_number(big.mark = ",")
  ) +
  scale_color_manual(
    values = c(
      "International — All years" = "#4E79A7",
      "International — Without COVID-19" = "#86BCDB",
      "International — COVID-19 only" = "#2F5D8A",
      "Domestic — All years" = "#F28E2B",
      "Domestic — Without COVID-19" = "#FFBE7D",
      "Domestic — COVID-19 only" = "#E86A17"
    )
  ) +
  labs(
    title = "Monthly tourism patterns in Munich",
    subtitle = "Hover for details; click legend entries to hide or show series",
    x = "Month",
    y = "Average tourism activity",
    color = "Origin and period"
  ) +
  theme_minimal(base_size = 10) +
  theme(
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold", size = 10),
    axis.title = element_text(size = 10),
    axis.text = element_text(size = 8),
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1),
    legend.position = "bottom",
    legend.title = element_text(size = 9),
    legend.text = element_text(size = 8),
    plot.title = element_text(size = 13),
    plot.subtitle = element_text(size = 9),
    plot.title.position = "plot",
    plot.margin = margin(8, 8, 8, 8),
    panel.spacing.x = grid::unit(0.3, "lines")
  )

interactive_plot <- ggplotly(
  covid_comparison_plot,
  tooltip = "text",
  dynamicTicks = FALSE,
  height = 520
)

interactive_plot <- plotly::layout(
  interactive_plot,
  autosize = TRUE,
  hovermode = "closest",
  margin = list(l = 55, r = 10, b = 135, t = 70),
  legend = list(
    orientation = "h",
    x = 0.5,
    xanchor = "center",
    y = -0.25,
    yanchor = "top",
    font = list(size = 10),
    groupclick = "togglegroup",
    itemclick = "toggle",
    itemdoubleclick = "toggleothers"
  )
)

interactive_plot <- plotly::config(
  interactive_plot,
  displaylogo = FALSE,
  responsive = TRUE
)

interactive_plot
```

## Highest and lowest months

The exact highest and lowest months are calculated from `seasonal_data_no_covid`. Using the non-COVID data prevents the unusually low values in 2020 and 2021 from determining what is described as Munich's normal low season.

```{r highest-lowest-months-rq1}
seasonal_extremes <- seasonal_data_no_covid %>%
  group_by(MONATSZAHL, AUSPRAEGUNG) %>%
  summarise(
    highest_month = month.name[month[which.max(mean_wert)]],
    highest_value = round(max(mean_wert)),
    lowest_month = month.name[month[which.min(mean_wert)]],
    lowest_value = round(min(mean_wert)),
    .groups = "drop"
  ) %>%
  transmute(
    measure = recode(
      MONATSZAHL,
      "Gäste" = "Guests",
      "Übernachtungen" = "Overnight stays"
    ),
    origin = recode(
      AUSPRAEGUNG,
      "Inland" = "Domestic",
      "Ausland" = "International"
    ),
    highest_month,
    highest_value,
    lowest_month,
    lowest_value
  )

seasonal_extremes %>%
  gt() %>%
  tab_header(
    title = "Highest and lowest tourism months",
    subtitle = "Monthly averages excluding 2020 and 2021"
  ) %>%
  cols_label(
    measure = "Tourism measure",
    origin = "Visitor origin",
    highest_month = "Highest month",
    highest_value = "Highest average",
    lowest_month = "Lowest month",
    lowest_value = "Lowest average"
  ) %>%
  fmt_number(
    columns = c(highest_value, lowest_value),
    decimals = 0,
    use_seps = TRUE
  ) %>%
  tab_options(
    table.font.size = px(14),
    heading.title.font.size = px(18),
    heading.subtitle.font.size = px(13),
    column_labels.font.weight = "bold",
    row.striping.include_table_body = TRUE
  )
```

## Analysis and answer

Munich tourism follows a clear annual cycle. Activity is lowest in winter, increases through spring, and is strongest from summer into early autumn before declining again. The table gives the exact peak and low month for each series.

International tourism is more seasonal than domestic tourism. International guest arrivals and overnight stays follow a similar pattern and both peak in September. Domestic tourism is spread more broadly across the high season: domestic overnight stays peak in July, while domestic guest arrivals peak in October. This difference shows that the month with the most arriving domestic visitors is not necessarily the month generating the most accommodation nights.

The September–October peak also coincides with Oktoberfest, which is held in Munich from late September into early October. The event is therefore a plausible catalyst for increased tourism during these months, particularly for the September international peak and the October domestic-arrival peak. However, the dataset contains no event indicator, so this timing suggests an association rather than proving that Oktoberfest caused the peak.

The COVID-only lines are substantially below the usual pattern, especially during spring and particularly for international tourism. Including 2020 and 2021 therefore lowers the all-years averages, although the overall seasonal shape remains visible. The non-COVID lines provide the clearest representation of Munich's usual seasonality.

Overall, **January and February form Munich's main low season, while July through October form its high season**. September is the clearest international peak; domestic tourism has a broader and less synchronized peak.

These results describe averages across many years. They may hide changes in seasonality over time, and the data identify when tourism is high or low but do not establish why those patterns occur.

## Research Question 2

How did Munich tourism change during the COVID-19 period, and how did domestic and international tourism recover afterwards?

### Tourism recovery after COVID-19

The COVID-19 pandemic caused an abrupt disruption to Munich tourism. This section examines monthly visitor arrivals (*Gäste*) and overnight stays (*Übernachtungen*) from 2020 onwards, separated into domestic (*Inland*) and international (*Ausland*) tourism.

Visitor arrivals measure the number of guests checking into accommodation. Overnight stays measure the number of nights spent in accommodation. Considering both measures helps distinguish between changes in visitor volume and changes in accommodation demand.

In the following charts, the blue line represents domestic tourism and the orange line represents international tourism. The dashed grey and black lines are linear models that respectively show the fitted trend in international and domestic visitor arrivals over time. They summarise the overall direction of recovery, rather than the trend of either group individually.

#### Visitor arrivals

```{r}
#| echo: false
#| warning: false
#| message: false
source("scripts/rq2/recovery_trend_plot_visitors.r")

visitor_lm
```

The visitor-arrivals chart shows a sudden collapse in both domestic and international travel in spring 2020. Total arrivals fell from about 576,000 in January 2020 and 571,000 in February to approximately 295,000 in April.

International tourism was especially affected by travel restrictions. International arrivals fell from around 237,000 in January 2020 to fewer than 48,000 in April. Domestic arrivals also fell sharply, from around 335,000 to approximately 25,000 over the same period.

The reopening phase was led by domestic tourism. In 2021, Munich recorded around 2.24 million domestic arrivals, compared with approximately 865,000 international arrivals. International travel recovered much more strongly from 2022 onwards.

By 2025, Munich recorded approximately 5.45 million domestic arrivals and 3.84 million international arrivals, resulting in a combined total of about 9.29 million visitor arrivals. This was similar to the 2024 total of approximately 9.26 million arrivals.

#### Overnight stays

```{r}
#| echo: false
#| warning: false
#| message: false

source("scripts/rq2/recovery_trend_plot_overnight.r")

overnight_lm
```

The overnight-stays chart shows the same pandemic shock, but it measures total accommodation demand rather than only the number of arriving visitors. A guest staying multiple nights contributes multiple overnight stays.

Domestic overnight stays recovered relatively quickly. They rose from around 4.82 million in 2020 to 5.56 million in 2021, then increased to 9.52 million in 2022 and 10.56 million in 2023. By 2025, domestic visitors generated approximately 10.89 million overnight stays.

International overnight stays recovered later but increased strongly after 2021. They rose from approximately 2.39 million overnight stays in 2021 to 6.49 million in 2022, 8.11 million in 2023, and 8.74 million in 2025. The 2025 value was close to the pre-pandemic international total of approximately 8.76 million overnight stays in 2019.

### Interpretation

The recovery can be divided into three phases:

- **Collapse in 2020:** COVID-19 restrictions led to a sharp decline in visitor arrivals and overnight stays. The fall was especially severe for international tourism.
- **Domestic-led reopening in 2021:** Domestic tourism returned earlier and made up the larger share of Munich's tourism recovery, while international tourism remained comparatively weak.
- **Broad recovery from 2022:** Both domestic and international tourism grew substantially. Domestic tourism surpassed its pre-pandemic level, whereas international tourism recovered strongly but remained slightly below its 2019 arrival volume in 2025.

The monthly patterns also show clear seasonality. Visitor arrivals and overnight stays tend to be higher in spring and summer and lower during winter. Therefore, comparing a month with the same month in the previous year is more meaningful than interpreting every month-to-month change as a change in tourism demand.

The dashed linear trends should be interpreted with care. They confirm an overall upward movement since 2020, but they simplify a recovery that was not linear: tourism first collapsed, then reopened gradually, and later recovered more quickly. The monthly domestic and international series provide the stronger evidence for the different recovery paths.

### Key takeaway

Munich tourism recovered strongly after the COVID-19 shock. By 2025, the city recorded about 9.29 million visitor arrivals, close to the 2024 level. Domestic tourism recovered first and exceeded its pre-pandemic volume, while international tourism returned more gradually and was still slightly below its 2019 visitor-arrival level in 2025.
