---
title: "Project Proposal"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
author:
  - "Aleksandar Neshov"
  - "Dinh Marcus Nguyen"
  - "Alexander Toropov"
date: today
---

## Dataset

### Source and license

| Item    | Information                                                   |
|---------|---------------------------------------------------------------|
| Source  | <https://opendata.muenchen.de/dataset/monatszahlen-tourismus> |
| License | Datenlizenz Deutschland Namensnennung 2.0 (dl-by-de)          |

### Description

The dataset contains monthly counts of guests and overnight stays in Munich, split by domestic and international visitors.

Each row represents one tourism observation for a specific combination of tourism measure (`MONATSZAHL`), visitor origin (`AUSPRAEGUNG`), year (`JAHR`), and month or yearly total (`MONAT`).

### Key variables

| Variable | Meaning | Type |
|----|----|----|
| `MONATSZAHL` | Metric type: guests or overnight stays | chr |
| `AUSPRAEGUNG` | Visitor origin: domestic / international / total | chr |
| `JAHR` | Year of observation | dbl |
| `MONAT` | Month of observation, or `"Summe"` for yearly totals | chr |
| `WERT` | Tourism value, either monthly value or yearly total depending on `MONAT` | dbl |

### Dimensions

| Dimension | Value |
|-----------|------:|
| Rows      | 1,632 |
| Columns   |    10 |

### Identification variable

The dataset contains an `_id` column, which can be used as a row identifier.

For the substantive tourism observations, each row can also be identified by the combination of:

`MONATSZAHL + AUSPRAEGUNG + JAHR + MONAT`

This combination describes the tourism measure, origin category, year, and month or yearly total for each observation.

### Fit with project requirements

The dataset meets the project requirements because it has more than 50 rows, a numeric outcome variable (`WERT`), a numeric explanatory variable (`JAHR`), and categorical explanatory variables (`MONATSZAHL` and `AUSPRAEGUNG`). Furthermore, this dataset is particularly interesting for our questions because it provides high-resolution, monthly data perfectly spanning the pre-pandemic, lockdown, and post-pandemic recovery periods, allowing us to precisely visualize the localized impact of COVID-19 on Munich's tourism.

------------------------------------------------------------------------

## Research questions

### Research question 1

**How do monthly tourism patterns in Munich vary across the year, and which months show the highest and lowest tourism activity?**

**Claim type:** Descriptive / exploratory.\
We describe seasonal patterns and identify months with high or low tourism activity.

**Output:** We will produce a seasonal line plot to visualize the peak and low months for both domestic and international tourists.

### Research question 2

**How did Munich tourism change during the COVID-19 period, and how did domestic and international tourism recover afterwards?**

**Claim type:** Descriptive / exploratory.\
We describe visible changes during and after COVID-19, without making strong causal claims.

**Output:** We will produce a continuous time-series line plot with reference lines marking the start and end of the pandemic to describe visible changes without making strong causal claims.

------------------------------------------------------------------------

## Context and target audience

Our project is framed in a **journalistic / public-interest context**. We want to explore how tourism in Munich has changed over time, especially seasonal patterns, the COVID-19 period, and differences between domestic and international tourism.

The target audience is a general audience interested in Munich, such as local residents, students, journalists, city administration, and tourism-related businesses. The analysis should therefore be understandable without advanced statistical knowledge.

Suitable presentation formats include:

- a short data report
- a Quarto document with explanatory text and visualisations
- line charts showing trends over time
- seasonal plots comparing months
- summary tables comparing domestic and international tourism

------------------------------------------------------------------------

## Initial EDA

### Setup

```{r}
library(tidyverse)

tourism <- read_csv("data/raw/tourism.csv")
```

### Dataset structure

```{r}
glimpse(tourism)
```

### Summary statistics for key variables

```{r}
tourism %>%
  summarise(
    rows = n(),
    columns = ncol(tourism),
    min_year = min(JAHR),
    max_year = max(JAHR),
    missing_wert = sum(is.na(WERT)),
    min_wert = min(WERT, na.rm = TRUE),
    median_wert = median(WERT, na.rm = TRUE),
    mean_wert = mean(WERT, na.rm = TRUE),
    max_wert = max(WERT, na.rm = TRUE)
  )
```

The dataset contains 1,632 rows and 10 columns. The years range from 2006 to 2026. The main outcome variable `WERT` has 72 missing values. For non-missing values, `WERT` ranges from 4,785 to 19,689,217, with a median of 498,710 and a mean of about 1,014,664.

### Category counts

```{r}
tourism %>%
  count(MONATSZAHL)

tourism %>%
  count(AUSPRAEGUNG)
```

The variable `MONATSZAHL` contains two tourism measures: `Gäste` and `Übernachtungen`, each with 816 observations. The variable `AUSPRAEGUNG` contains three categories: `Ausland`, `Inland`, and `insgesamt`, each with 544 observations.

Here, `insgesamt` means the total of domestic and international tourism.

### Missing values

```{r}
tourism %>%
  summarise(across(everything(), ~ sum(is.na(.))))
```

Important missing values already visible in the dataset:

| Variable                         | Missing values |
|----------------------------------|---------------:|
| `WERT`                           |             72 |
| `VERAEND_VORMONAT_PROZENT`       |            192 |
| `VERAEND_VORJAHRESMONAT_PROZENT` |             72 |
| `ZWOELF_MONATE_MITTELWERT`       |            192 |

------------------------------------------------------------------------

## Plots

### Plot 1: Distribution of tourism values

```{r}
tourism %>%
  ggplot(aes(x = WERT)) +
  geom_histogram(bins = 30) +
  scale_x_log10() +
  labs(
    title = "Distribution of tourism values",
    x = "Tourism value WERT, log scale",
    y = "Number of observations"
  )
```

The distribution of `WERT` is strongly right-skewed. Most observations are much smaller than the largest values. The highest values likely come from yearly totals where `MONAT` is `"Summe"`, while regular monthly observations are lower. A log scale makes the distribution easier to inspect.

### Plot 2: Tourism trend over time

```{r}
tourism %>%
  filter(MONAT == "Summe") %>%
  filter(AUSPRAEGUNG %in% c("Inland", "Ausland")) %>%
  ggplot(aes(x = JAHR, y = WERT, color = AUSPRAEGUNG)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ MONATSZAHL, scales = "free_y") +
  labs(
    title = "Tourism in Munich over time",
    x = "Year",
    y = "Yearly tourism value",
    color = "Origin category"
  )
```

This plot gives a first overview of tourism trends over time. It compares domestic and international tourism and separates guests from overnight stays. The plot is relevant for our research questions about long-term trends and changes around the COVID-19 period.

For this plot, only rows where `MONAT == "Summe"` are used, because these rows represent yearly totals. This avoids mixing monthly observations with yearly totals.

### Plot 3: Monthly pattern

```{r}
tourism %>%
  filter(MONAT != "Summe") %>%
  filter(AUSPRAEGUNG %in% c("Inland", "Ausland")) %>%
  mutate(
    month = as.integer(str_sub(MONAT, 5, 6))
  ) %>%
  group_by(month, MONATSZAHL, AUSPRAEGUNG) %>%
  summarise(mean_wert = mean(WERT, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = month, y = mean_wert, color = AUSPRAEGUNG)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ MONATSZAHL, scales = "free_y") +
  scale_x_continuous(breaks = 1:12) +
  labs(
    title = "Average monthly tourism patterns in Munich",
    x = "Month",
    y = "Average tourism value",
    color = "Origin category"
  )
```

This plot helps us examine seasonal patterns. It shows average tourism values by month and compares domestic and international tourism. This is relevant for identifying months with particularly high or low tourism activity.

------------------------------------------------------------------------

## Notes on data quality issues

Several data quality issues are already visible in the initial exploration.

First, the variable `MONAT` contains both monthly values such as `"202401"` and yearly totals marked as `"Summe"`. These should not be mixed without care, because yearly totals are much larger than monthly observations.\
We will filter out the "Summe" rows when doing monthly analysis, and use `tsibble::yearmonth()` with a format argument to parse the strings into a proper time-series index.

Second, `WERT` has 72 missing values. These missing values appear especially relevant for 2026, which is likely incomplete at the time of data publication.\
We will filter out rows where `WERT` is `NA` prior to conducting our time-series analysis, effectively ending our dataset in December 2025.

Third, the variables `VERAEND_VORMONAT_PROZENT` and `ZWOELF_MONATE_MITTELWERT` have more missing values than `WERT`. This is expected because percentage changes and moving averages cannot always be calculated for early or incomplete observations.\
We must be careful to filter out "insgesamt" when grouping or comparing domestic and international tourism so we don't double-count visitors.

Fourth, `AUSPRAEGUNG` includes `"insgesamt"`, which is the total of `"Inland"` and `"Ausland"`. Therefore, `"insgesamt"` should not be treated as an independent third group when comparing domestic and international tourism. While this looks like an error, domain knowledge confirms this is an accurate reflection of the post-COVID-19 rebound (comparing normal months to pandemic lockdown months). No cleaning is required, but we must be aware of this extreme volatility if modeling.

------------------------------------------------------------------------

## Analysis plan

### Research question 1

**Question:**\
How do monthly tourism patterns in Munich vary across the year, and which months show the highest and lowest tourism activity?

#### Variables used

| Variable      | Meaning                                             |
|---------------|-----------------------------------------------------|
| `WERT`        | Tourism value (outcome variable)                    |
| `MONAT`       | Month / time variable (`yearmonth` object)          |
| `MONATSZAHL`  | Tourism measure: (`Gäste` or `Übernachtungen`)      |
| `AUSPRAEGUNG` | Origin category: (`Inland`, `Ausland`, `insgesamt`) |

#### Analytical approach

We will produce a seasonal line plot using `ggplot2` mapping the extracted month (1-12) to the x-axis and the average `WERT` to the y-axis, grouped and colored by `AUSPRAEGUNG`. This plot will directly answer our research question by visually identifying the peak and low seasons and showing if domestic and international tourists follow the same seasonal calendar.

#### Cleaning / transformation steps

We will use `dplyr::filter(MONAT != "Summe")` to remove annual totals and `filter(!is.na(WERT))` to drop the 72 missing, unrecorded values for 2026. We will convert `MONAT` to a `tsibble::yearmonth` object. Finally, we will use `lubridate::month()` to extract the numerical month from the `MONAT` object so we can aggregate the data across all years to find the average seasonal trend. We will filter out `AUSPRAEGUNG == "insgesamt"` to avoid double-counting.

#### Uncertainties / risks

Averaging data across all years (2006-2025) might obscure changing trends over time (e.g., if Oktoberfest became more popular recently than in 2006). The COVID-19 years (2020-2021) also heavily distort the "average" monthly values, so we may need to exclude pandemic years when calculating the typical seasonal baseline.

------------------------------------------------------------------------

### Research question 2

**Question:**\
How did Munich tourism change during the COVID-19 period, and how did domestic and international tourism recover afterwards?

#### Variables used

| Variable      | Meaning                            |
|---------------|------------------------------------|
| `WERT`        | Tourism value (outcome variable)   |
| `MONAT`       | Time variable (`yearmonth` object) |
| `MONATSZAHL`  | `Gäste` or `Übernachtungen`        |
| `AUSPRAEGUNG` | `Inland`, `Ausland`, `insgesamt`   |

#### Analytical approach

We will produce a continuous time-series line plot with `ggplot2` mapping the `yearmonth` `MONAT` to the x-axis and `WERT` to the y-axis, colored by `AUSPRAEGUNG`. We will add vertical reference lines (`geom_vline`) in March 2020 (start of lockdowns) and early 2022 (lifting of restrictions) to visually frame the pandemic period. This will answer the research question by showing the immediate drop and tracking the slope of the recovery curves for both visitor types.

#### Cleaning / transformation steps

As with RQ1, we will filter out `"Summe"` rows and missing `WERT` rows, and filter out the `"insgesamt"` category to focus strictly on comparing domestic and international trends.

#### Uncertainties / risks

The main limitation is that the dataset alone cannot prove that COVID-19 caused the observed changes, though the timing is highly suggestive. Furthermore, comparing recovery rates between domestic and international tourism may be confounded by differing international border policies (e.g., travel bans) which are not captured in this dataset.

------------------------------------------------------------------------

## Workflow organisation

### Code style

We will follow the tidyverse style guide to keep the code readable and consistent.

Main conventions:

- use `snake_case` for object and variable names
- use clear and meaningful names
- use the tidyverse pipe `%>%`
- keep code chunks short and focused
- add comments only where they explain non-obvious steps

To enforce this, we will use a documented convention in the project README. We may also use `styler` to format code automatically:

```{r, eval=FALSE}
install.packages("styler")
styler::style_file("proposal.qmd")
```

### Packages

We plan to use the following R packages:

| Package | Purpose |
|----|----|
| `tidyverse` | Data cleaning, summaries, and visualisation |
| `readr` | Reading CSV files |
| `dplyr` | Data manipulation |
| `ggplot2` | Plots |
| `lubridate` | Date and month handling |
| `janitor` | Cleaning column names and simple data cleaning in later project stages |
| `broom` | Tidying model outputs, if we use a model |

We may use `renv` to make the project reproducible across different computers:

```{r, eval=FALSE}
install.packages("renv")
renv::init()
renv::snapshot()
```

This would lock the package versions used in the project.

### Git workflow

We will use a simple task-based Git workflow.

The `main` branch will contain the stable version of the project. Each group member will work on a separate branch or on a task-specific branch.

Possible branches:

| Branch           | Purpose                              |
|------------------|--------------------------------------|
| `main`           | Stable project version               |
| `proposal`       | Work on project proposal             |
| `data-cleaning`  | Data cleaning scripts                |
| `visualisations` | Plots and exploratory visualisations |
| `analysis-plan`  | Analysis planning and methods        |
| `final-report`   | Final report preparation             |

Before merging into `main`, we will check that the Quarto document still renders correctly and that the code runs without errors.

### `.gitignore`

Our `.gitignore` file was generated from the standard templates for R, Python, Windows, macOS, and Linux, and extended with Quarto-specific entries.

We exclude operating-system files such as `.DS_Store`, `Thumbs.db`, and Linux temporary files. We also exclude R session files such as `.Rhistory`, `.RData`, and `.Rproj.user/`, because they are user-specific and should not be shared through Git.

For reproducibility and security, we exclude local environment files such as `.env`, `.Renviron`, virtual environments, cache folders, and temporary build files. Since the project uses Quarto, we also exclude generated Quarto folders and cache files such as `/.quarto/`, `/_site/`, `/*_files/`, and `/*_cache/`.

Rendered output files and cached intermediate files should be regenerated from the Quarto source document instead of being manually tracked in Git.
